---
title: "charts"
author: "Katica Ristic"
date: "July 6, 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Charts 

Charts made with a help of the link:

<https://moderndata.plot.ly/interactive-r-visualizations-with-d3-ggplot2-rstudio/>.



```{r DATAload, echo = FALSE, warning = FALSE, cache = FALSE, message = FALSE}

library(plotly)
library(shiny)

setwd("D:/MathJob/Tijana")
data <- read.csv("PacijentiCSV.csv")
attach(data)

```

## eGFR -  male vs female

```{r boxes eGFR MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, color = ~MF, type = "box" )
p


```

## eGFR - patient vs control group


```{r boxes eGFR PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, color = ~PatientControl, type = "box" )
p


```

## ADCcortex - male vs female

```{r boxes ADCcortex MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCcortex, color = ~MF, type = "box" )
p


```

## ADCcortex - patient vs control group

```{r boxes ADCcortex PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCcortex, color = ~PatientControl, type = "box" )
p


```

## ADCmedulla - male vs female


```{r boxes ADCmedulla MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCmedulla, color = ~MF, type = "box" )
p


```

## ADCmedulla - patient vs control group

```{r boxes ADCmedulla PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCmedulla, color = ~PatientControl, type = "box" )
p


```

## ADCwholekidney - male vs female

```{r boxes ADCwholekidney MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCwholekidney, color = ~MF, type = "box" )
p


```

## ADCwholekidney - patient vs control group

```{r boxes ADCwholekidney PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCwholekidney, color = ~PatientControl, type = "box" )
p


```

## ADCCMD - male vs female

```{r boxes ADCCMD MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCCMD, color = ~MF, type = "box" )
p


```

## ADCCMD - patient vs control group

```{r boxes ADCCMD PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~ADCCMD, color = ~PatientControl, type = "box" )
p


```

## Chart eGFR vs ADCcortex 


```{r Chart eGFR vs ADCcortex for MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCcortex , 
             text = paste("Gender: ", MF, '\n' ,"Group: ", PatientControl  ),
             mode = "markers", 
             color = ~MF, 
             size = ADCcortex,
             symbol = ~PatientControl)
p


```


## Chart eGFR vs ADCmedulla  


```{r Chart eGFR vs ADCmedulla for MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCmedulla , 
             text = paste("Gender: ", MF, '\n' ,"Group: ", PatientControl), 
             mode = "markers", 
             color = ~MF, 
             size = ADCmedulla,
             symbol = ~PatientControl )
p


```



## Chart eGFR vs ADCwholekidney 


```{r Chart eGFR vs ADCwholekidney for MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCwholekidney , 
             text = paste("Gender: ", MF, '\n' ,"Group: ", PatientControl), 
             mode = "markers", 
             color = ~MF, 
             size = ADCwholekidney,
             symbol = ~PatientControl )
p


```


## Chart eGFR vs ADCCMD 


```{r Chart eGFR vs ADCCMD for MF, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCCMD , 
             text = paste("Gender: ", MF, '\n' ,"Group: ", PatientControl), 
             mode = "markers", 
             color = ~MF, 
             size = ADCCMD,
             symbol = ~PatientControl )
p


```
   
   
## Chart eGFR vs ADCcortex for patient and control group


```{r Chart eGFR vs ADCcortex for PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCcortex , 
             text = paste("Group: ", PatientControl  ), 
             mode = "markers", 
             color = ~PatientControl, 
             size = ADCcortex )
p


```





## Chart eGFR vs ADCmedulla for patient and control group


```{r Chart eGFR vs ADCmedulla for PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCmedulla , 
             text = paste("Gender: ", PatientControl), 
             mode = "markers", 
             color = ~PatientControl, 
             size = ADCmedulla )
p


```



## Chart eGFR vs ADCwholekidney for patient and control group 


```{r Chart eGFR vs ADCwholekidney for PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCwholekidney , 
             text = paste("Gender: ", PatientControl), 
             mode = "markers", 
             color = ~PatientControl, 
             size = ADCwholekidney )
p


```


## Chart eGFR vs ADCCMD for patient and control group


```{r Chart eGFR vs ADCCMD for PC, echo=FALSE, warning = FALSE, cache = FALSE, message = FALSE}

p <- plot_ly(data, x = ~eGFR, y = ADCCMD , 
             text = paste("Gender: ", PatientControl), 
             mode = "markers", 
             color = ~PatientControl, 
             size = ADCCMD )
p


```
   
   
   
   
   
               
   
   
             